Dynamics of isoflurane-induced vasodilation and blood flow of cerebral vasculature revealed by multi-exposure speckle imaging

### 2021-08-18 article.tex
Revisions

\title{Dynamics of isoflurane-induced vasodilation and blood flow of cerebral vasculature revealed by laser speckle contrast imaging}

@@

\abstractsection{Background}
Anesthetized animal models are used extensively during neurophysiological and behavioral studies despite systemic effects from anesthesia that undermine both accurate interpretation and translation to awake human physiology. The majority of work examining the impact of anesthesia on cerebral blood flow (CBF) has been limited to before and after measurements with limited spatial resolution.

@@

\abstractsection{Results}
The large anatomical changes caused by isoflurane are depicted with wide-field imagery and video highlighting the induction of anesthesia. Compared to the awake state, we measured, on average, a 14.1\% increase in surface vessel diameter accompanied by a 96.0\% increase in vascular flux and 84.7\% increase in parenchyma tissue perfusion. We also examined the dynamics of anesthesia induction and found that blood flow increased faster in arterioles than in venules or parenchyma regions.

\abstractsection{Comparison with Existing Methods}
MESI offers robust hemodynamic measurements across large fields-of-view and high temporal resolutions sufficient for continuous visualization of cerebrovascular events.

@@

% Include a list of up to six keywords after the abstract
\begin{keyword}
    laser speckle contrast imaging \sep
    multi-exposure speckle imaging \sep
    awake imaging \sep
    anesthesia \sep
    hemodynamics \sep
    cerebral blood flow
\end{keyword}

@@

The 15 raw intensity images were converted to speckle contrast images ($K = \sigma_{s} / {\langle{I}\rangle}$) with a 7$\times$7-pixel sliding window and used to calculate an estimate of the speckle correlation time ($\tau_c$) at each pixel with the multi-exposure speckle visibility expression \cite{Parthasarathy:2008el}:
%
% Equation 1 - MESI Equation
\begin{equation}
    \label{eq:mesi}
    K(T,\tau_c) =
        \left(
        \beta\rho^2\frac{e^{-2x} - 1 + 2x}{2x^2} +
        4\beta\rho(1 - \rho)\frac{e^{-x} - 1 + x}{x^2} +
        \nu_{noise}
        \right)^{1/2}
\end{equation}

\noindent where $T$ is the camera exposure time, $x = T/\tau_c$, $\rho$ is the fraction of light that is dynamically scattered, $\beta$ is a normalization factor that accounts for speckle sampling, and $\nu_{noise}$ represents exposure-independent instrument noise and nonergodic variances. Eq.~(\ref{eq:mesi}) was fitted with the Levenberg-Marquardt nonlinear least squares algorithm \cite{Lourakis:J2fCMU5i} using a custom program written in C.

Because $\tau_c$ is inversely related to the speed of the moving scatterers \cite{Bonner:1981hg,Briers:1996kfa}, the inverse correlation time ($ICT = 1/\tau_c$) is frequently used as a metric for quantifying blood flow in vasculature and perfusion in parenchyma \cite{Ayata:2004ba,Strong:2005kj,Sullender:2018ff}. Recent work to improve the quantitative accuracy of MESI flow measurements in vasculature takes into account the presence of multiple dynamic scattering events instead of assuming only a single dynamic scattering event per photon \cite{Kazmi:2015du}. This is achieved by scaling the fitted $ICT$ value by the diameter of the vessel in order to obtain an estimate of ``vascular flux'' that better accounts for variations in vascular volume sampling \cite{Kazmi:2015du,Richards:2017df}. This paper uses the vascular flux metric (arbitrary units) for all vessel measurements and $ICT$ (s$^{-1}$) for all parenchymal perfusion measurements.

@@

Cranial window implanted mice were head constrained on the awake imaging system and allowed to acclimate for 5-10 minutes. The awake hemodynamic baseline was then defined prior to continuous imaging by acquiring 50 MESI frames, averaging the speckle contrast across each exposure time, and computing the corresponding $ICT$ frame using Eq.~(\ref{eq:mesi}). A custom 3D-printed inhalation nose-cone connected to the anesthesia vaporizer and scavenger was then placed several millimeters away from the subject and used to administer medical air (0.5 L/min) with no isoflurane (0\%). Continuous MESI was initiated ($t = 0$) and used to monitor the awake subject for 10 minutes before increasing isoflurane to 2.0\% to induce anesthesia. After one minute of exposure ($t = 11$ minutes), the nose-cone was fully positioned over the subject to ensure targeted delivery of isoflurane to the subject. After two minutes of exposure ($t = 12$ minutes), a feedback heating pad (55-7030, Harvard Apparatus, Inc.) was placed beneath the subject to maintain a 37 $^\circ$C body temperature. The now anesthetized subject was continuously monitored with MESI until $t = 30$ minutes, at which point the anesthetized hemodynamic state was defined by acquiring and averaging an additional 50 MESI frames. The subject was then removed from the imaging system and allowed to recover from anesthesia on a heating pad before being transferred back into its cage. Each subject underwent three imaging sessions with at least one full day of recovery between each session.

@@

All data processing was performed using MATLAB (R2021a, MathWorks, Inc.). While head fixation can greatly minimize locomotion-associated brain movement \cite{Dombeck:2007gr}, it does not completely eliminate motion artifacts. The induction of general anesthesia also causes major changes in posture that can result in lateral displacements of the brain relative to the skull. In order to account for these shifts, the open source image registration toolbox \texttt{Elastix} \cite{Klein:2010gr} was used to rigidly align all speckle contrast frames to the baseline awake imagery prior to fitting Eq.~(\ref{eq:mesi}). The resulting aligned MESI $ICT$ frames were then smoothed temporally with a central moving average filter ($k=5$) and rendered to video at 20 frames-per-second (8$\times$ speed). $ICT$ timecourses spanning the entire imaging session were then extracted from the aligned data from both vascular and parenchymal regions of interest (ROIs).

\subsubsection{Calculating Vascular Flux}
In order to calculate the vascular flux \cite{Kazmi:2015du}, the diameter of vessels ($d_{vessel}$) were estimated using the 5 ms exposure speckle contrast frames. Cross-sectional profiles were extracted across each vessel of interest at every timepoint and fitted to a Gaussian function in order to calculate the full-width at half-maximum (FWHM) of the distribution over time. Because LSCI is most sensitive to moving scatterers, i.e. erythrocytes, this method of estimating $d_{vessel}$ is a measure of the inner diameter of the vessel. The vascular flux was then computed by scaling the MESI $ICT$ by $d_{vessel}$ at each timepoint (arbitrary units). The percent change in vessel diameter and vascular flux were calculated using the average of the entire awake state for the baseline.

\subsubsection{Calculating Blood Flow Rise Time}
To examine the dynamics of isoflurane-induced flow changes, the rise time in measured blood flow was calculated. Rise time was defined as the time elapsed following the start of anesthesia for $ICT$ values to increase by 50\% relative to the awake baseline. A spatially-resolved blood flow rise time map was generated for Subject 1 by performing this calculation pixel-wise across the entire field-of-view. Analogous rise time calculations were performed on the vascular flux and parenchymal tissue perfusion metrics across all subjects and imaging sessions using the previously extracted ROI timecourses.

\subsubsection{Statistical Analysis}
All statistical analyses were performed using MATLAB (R2021a, MathWorks, Inc.). The differences between arteriole, venule, and parenchymal measurements were evaluated with one-way ANOVAs using the Bonferroni method for multiple comparisons. $p \leq 0.05$ was considered statistically significant.

@@

The systematic changes caused by general anesthesia with isoflurane in Subject 1 are shown in Fig.~\ref{fig:system}b and Video 1, which highlights the transition from the awake to the anesthetized state. The visible decrease in speckle contrast and corresponding increase in MESI $ICT$ are representative of an increase in CBF caused by the vasodilatory effects of isoflurane. Dilation is seen across much of the surface vasculature, in particular the large central vein and numerous smaller vessels that have become more prominent in the anesthetized imagery as blood flow increased. These large alterations to the vasculature complicated rigid image registration and precluded the generation of relative flow change imagery.

% Figure 1 - System Schematic + Awake vs. Anesthetized Imagery
\begin{figure}
    \includegraphics[width=\textwidth]{Figure1.pdf}
    \caption {
        (a) MESI system schematic with 2$\times$ magnification: A (Standard Iris, ID15, Thorlabs, Inc.), M ({{\O}}1$"$ Broadband Dielectric Mirror, BB1-E02, Thorlabs, Inc.), L1 ($f =$ 50 mm, Steinheil Triplet Achromatic, 67-422, Edmund Optics, Inc.), F (685 $\pm$ 20 nm bandpass filter, S685/40m, Chroma Technology Corp.), and L2 ($f = 100$ mm, Achromatic Doublet, AC254-100-A, Thorlabs, Inc.). (b) Single exposure LSCI (5 ms) and MESI inverse correlation time ($ICT$) during awake and anesthetized states in Subject 1 (Scale bar = 1 mm) (Video 1, MPEG-4, 35.8 MB).
    }
    \label{fig:system}
\end{figure}

The temporal dynamics of the awake-to-anesthetized transition can be seen in Fig.~\ref{fig:timecourse}, where the change in blood flow within two veins (R1 and R2) and one parenchyma region (R3) were analyzed during the extended two-hour imaging session in Subject 1. The cross sections used to estimate $d_{vessel}$ are depicted in Fig.~\ref{fig:timecourse}a with the resulting changes in vessel diameter shown in Fig.~\ref{fig:timecourse}b. The width of the two vessels remained steady throughout the entirety of the awake measurements but increased under anesthesia, with the larger vessel (R1) growing by over 50\% after an hour of exposure. This vasodilation was also accompanied by an increase in blood flow as measured by the percent change in vascular flux and parenchymal tissue perfusion as shown in Fig.~\ref{fig:timecourse}c. The vascular flux in R1 and R2 increased by 270\% and 144\%, respectively, while the tissue perfusion in R3 grew by 141\%. The large deviation shortly after $t = 80$ minutes was caused by the subject briefly experiencing breathing difficulties that required the anesthesia be reduced to 1.5\% isoflurane for several minutes. The breathing abnormality and change in anesthesia dosage had no apparent effects on the vessel size. Repeated measurements in the subject (Figs.~\ref{fig:timecourse}d-f) illustrate the stability of day-to-day MESI estimates of blood flow as well as the consistency of the hemodynamic response to isoflurane anesthesia.

@@

The percent changes in vessel diameter and blood flow between the awake and anesthetized states were aggregated across all subjects and imaging sessions as shown in Fig.~\ref{fig:aggregated}. The anesthetized measurements for Subject 1, which underwent the extending imaging sessions, were taken at $t \approx 80$ minutes to match the timing of the other subjects (20 minutes after the start of anesthesia). A total of 108 measurements were included in the analysis (33 arteriole, 29 venule, 46 parenchyma). Similar to the results seen in Fig.~\ref{fig:timecourse}, the vasodilatory effects of isoflurane induced large changes in the cortical hemodynamics. On average, vessel width increased by 14.1\%, with arterioles increasing by 15.0\% and venules increasing by 13.0\% (Fig.~\ref{fig:aggregated}a). There was no significant difference in the percent change in width between arterioles and venules ($p = 0.61$). Vascular flux increased, on average, by 96.0\%, with arterioles increasing by 81.7\% and venules increasing by 112.2\% (Fig.~\ref{fig:aggregated}b). Parenchymal tissue perfusion increased by 84.7\%. While the ANOVA was significant ($p = 0.041$), the multiple pairwise comparisons between arterioles, venules, and parenchyma were all non-significant ($p > 0.05$).

% Figure 3 - Aggregated Awake to Anesthetized Results
\begin{figure}
    \includegraphics[width=\textwidth]{Figure3.pdf}
    \caption {
        (a) Percent change in vessel diameter for arterioles and venules between awake and anesthetized states ($p > 0.05$). (b) Percent change in arteriolar and venous vascular flux and parenchymal tissue perfusion between awake and anesthetized states ($p > 0.05$). Measurements in both charts aggregated across all subjects and all imaging sessions.
    }
    \label{fig:aggregated}
\end{figure}

The dynamics of the isoflurane-induced flow changes in Subject 1 can be seen in the blood flow rise time map in Fig.~\ref{fig:risetime}a. $ICT$ values across the entire field-of-view increased by at least 50\% in just over one minute following the start of anesthesia. The vasculature, including both arterioles and venules, appears to change flow faster than the parenchyma. However, upon aggregating ROI measurements across all subjects and imaging sessions (Fig.~\ref{fig:risetime}b), arterioles have significantly faster rise times ($40.5$ seconds) compared to venules ($50.7$ seconds, $p = \num{3.7e-6}$) or parenchyma ($46.1$ seconds, $p = 0.006$). Parenchymal regions also have significantly faster rise times than venules ($p = 0.043$).

% Figure 4 - Hemodynamic Rise Time
\begin{figure}
    \includegraphics[width=\textwidth]{Figure4.pdf}
    \caption {
        (a) Blood flow rise time map depicting the time elapsed following the start of anesthesia for the inverse correlation time ($ICT$) to increase by 50\% relative to awake baseline in Subject 1 (Scale bar = 1 mm). (b) Time for arteriolar and venous vascular flux and parenchymal tissue perfusion to increase by 50\% relative to the awake baseline (***$p = \num{3.7e-6}$, **$p = 0.006$, *$p = 0.043$). Measurements aggregated across all subjects and all imaging sessions.
    }
    \label{fig:risetime}
\end{figure}

@@

The systemic effects of isoflurane undermine the reliability of neurophysiological and behavioral studies because anesthesia is a fundamentally unnatural physiological state \cite{Gao:2017tw}. Imaging blood flow on the cortical surface with LSCI during the induction of anesthesia directly visualizes and quantifies these large hemodynamic changes. On average, the vasodilation caused by isoflurane increased vessel diameter by 14.1\% across all trials, with arterioles increasing by 15.0\% and venules increasing by 13.0\% (Fig.~\ref{fig:aggregated}a). Direct microscope measurements in fentanyl- and nitrous oxide-anesthetized rats documented a 17\% increase in arteriolar diameter and 6\% increase in venule diameter after the topical application of isoflurane \cite{Koenig:1994rn}. Similar measurements in pentobarbital-anesthetized dogs observed 10-28\% increases in the diameter of small arterioles following the inhalation of isoflurane \cite{Iida:1998th}. Both studies found that isoflurane dilated arterioles in a concentration-dependent manner. More recent measurements in mice using optical coherence tomography (OCT) angiography found that isoflurane increased artery diameter by 12-55\% and vein diameter by 14-22\%, depending on vessel branch order \cite{Rakymzhan.2021}.

Much larger changes were measured in surface vessel blood flow, with vascular flux increasing on average by 96.0\% across all trials (Fig.~\ref{fig:aggregated}b). This suggests that increases in blood flow velocity rather than vasodilation are largely responsible for the measured change in flow. Because few imaging modalities are capable of measuring the dynamics of blood flow with sufficient spatial resolution to distinguish individual vessels, there has been limited work on this topic in awake animals. One study using laser Doppler flowmetry found only non-significant 18\% increases in both CBF and red blood cell (RBC) velocity in the barrel cortex after mice were anesthetized with isoflurane \cite{Takuwa:2012ee}. However, because large blood vessels were avoided, the laser Doppler measurements only sampled smaller subsurface vasculature that may exhibit different responses to isoflurane than the larger surface vasculature imaged in this study. The vascular flux metric calculated from the speckle correlation time is also not analogous to either the CBF or RBC velocity measured by laser Doppler flowmetry, so the results are not directly comparable \cite{Kazmi:2015du}. Another more recent study using Doppler OCT measured a 55\% increase in total (volumetric) blood flow in mice under isoflurane \cite{Rakymzhan.2021}. Unlike the results presented above, this change was largely driven by vasodilation rather than an increase in blood flow velocity.

Within the parenchyma, tissue perfusion increased on average by 84.7\% across all trials after isoflurane exposure (Fig.~\ref{fig:aggregated}b). The narrower distribution of values compared to the vascular measurements is likely a byproduct of the more homogenous structure of the unresolvable capillaries within the parenchyma. fMRI measurements in isoflurane-anesthetized rats have reported $\sim$50\% increases in global CBF with the cerebral cortex experiencing the smallest regional increase (20\%) compared to the awake state \cite{Sicard:2003dj}. However, these numbers are not directly comparable because fMRI samples a much larger volume of the brain than LSCI, which has both a smaller field-of-view and only penetrates several hundred microns into the cortex.

While LSCI has been utilized extensively with awake imaging \cite{Takuwa:2011jr,Seto:2014ga,Lu:2017bo,Balbi:2017cj,Sunil:2020ac}, this is one of the first uses of MESI in an awake animal model \cite{He:2020}. Traditional single-exposure LSCI would likely have underestimated the magnitude of the flow changes caused by isoflurane because individual exposure times are only sensitive to a small range of flow rates \cite{Parthasarathy:2008el}. Failing to account for the increase in multiple dynamic scattering events caused by vasodilation would also have resulted in an underestimation of the flow changes in the surface vasculature. Using the raw $ICT$ value instead of the diameter-scaled vascular flux metric in Fig.~\ref{fig:timecourse}c would have measured only a 141\% increase in the large vessel instead of 270\%. The repeatability of the day-to-day measurements shown in Figs.~\ref{fig:timecourse}d-f further demonstrate the robustness of MESI as a chronic blood flow imaging technique \cite{Kazmi:2013hp}, even when observing a fully-conscious, moving animal.

@@

The effects of isoflurane fundamentally undermine the imaging of cortical hemodynamics in the normal physiological state. Blood flow imaging techniques such as LSCI, laser Doppler flowmetry, and fMRI require greater sensitivity because of the suppression of the hemodynamic response \cite{Takuwa:2012ee} and measure a delayed neurovascular coupling \cite{Pisauro:2013cx}. Anatomical imaging with two-photon microscopy only captures heavily-dilated vasculature unrepresentative of the awake resting state \cite{Lyons:2016bd}. The consistency and reliability of all imaging techniques can suffer from the effects of prolonged isoflurane exposure, which can cause continual vasodilation as seen in Fig.~\ref{fig:timecourse}b. This may result in extended imaging sessions documenting disparate anatomies and physiologies between the beginning and conclusion of an experiment.

>>>>

\title{Dynamics of isoflurane-induced vasodilation and blood flow of cerebral vasculature revealed by multi-exposure speckle imaging}

@@

\abstractsection{Background}
Anesthetized animal models are used extensively during neurophysiological and behavioral studies despite systemic effects from anesthesia that undermine both accurate interpretation and translation to awake human physiology. The majority of work examining the impact of anesthesia on cerebral blood flow (CBF) has been restricted to before and after measurements with limited spatial resolution.

@@

\abstractsection{Results}
The large anatomical changes caused by isoflurane are depicted with wide-field imagery and video highlighting the induction of general anesthesia. Within minutes of exposure, both vessel diameter and blood flow increased drastically compared to the awake state and remained elevated for the duration of imaging. An examination of the dynamics of anesthesia induction reveals that blood flow increased faster in arteries than in veins or parenchyma regions.

\abstractsection{Comparison with Existing Methods}
MESI offers robust hemodynamic measurements across large fields-of-view and high temporal resolutions sufficient for continuous visualization of cerebrovascular events featuring major changes in blood flow.

@@

% Include a list of up to six keywords after the abstract
\begin{keyword}
    multi-exposure speckle imaging \sep
    laser speckle contrast imaging \sep
    awake imaging \sep
    anesthesia \sep
    hemodynamics \sep
    cerebral blood flow
\end{keyword}

@@

The 15 raw intensity images were converted to speckle contrast images ($K = \sigma_{s} / {\langle{I}\rangle}$) with a 7$\times$7-pixel sliding window and used to calculate an estimate of the speckle correlation time ($\tau_c$) at each pixel with the multi-exposure speckle visibility expression \cite{Parthasarathy:2008el}:
%
% Equation 1 - MESI Equation
\begin{equation}
    \label{eq:mesi}
    K(T,\tau_c) =
        \left(
        \beta\rho^2\frac{e^{-2x} - 1 + 2x}{2x^2} +
        4\beta\rho(1 - \rho)\frac{e^{-x} - 1 + x}{x^2} +
        \nu
        \right)^{1/2}
\end{equation}

\noindent where $T$ is the camera exposure time, $x = T/\tau_c$, $\rho$ is the fraction of light that is dynamically scattered, $\beta$ is a normalization factor that accounts for speckle sampling, and $\nu$ represents exposure-independent instrument noise and nonergodic variances. Eq.~(\ref{eq:mesi}) was fitted with the Levenberg-Marquardt nonlinear least squares algorithm \cite{Lourakis:J2fCMU5i} using a custom program written in C.

Because $\tau_c$ is inversely related to the speed of the moving scatterers \cite{Bonner:1981hg,Briers:1996kfa}, the inverse correlation time ($ICT = 1/\tau_c$) is frequently used as a metric for quantifying blood flow in vasculature and perfusion in parenchyma \cite{Ayata:2004ba,Strong:2005kj,Sullender:2018ff}. Recent work to improve the quantitative accuracy of MESI flow measurements in vasculature takes into account the presence of multiple dynamic scattering events instead of assuming only a single dynamic scattering event per photon \cite{Kazmi:2015du}. This is achieved by scaling the fitted $ICT$ value by the diameter of the vessel in order to obtain an estimate of ``vascular flux'' that better accounts for variations in vascular volume sampling \cite{Kazmi:2015du,Richards:2017df}. This paper uses the vascular flux metric (arbitrary units) for all vessel measurements and $ICT$ (s$^{-1}$) for all parenchyma perfusion measurements.

@@

Cranial window implanted mice were head constrained on the awake imaging system and allowed to acclimate for 5-10 minutes. The awake hemodynamic baseline was then defined prior to continuous imaging by acquiring 50 MESI frames, averaging the speckle contrast across each exposure time, and computing the corresponding $ICT$ frame using Eq.~(\ref{eq:mesi}). A custom 3D-printed inhalation nose-cone connected to the anesthesia vaporizer and scavenger was then placed several millimeters away from the subject and used to administer medical air (0.5 L/min) with no isoflurane (0\%). Continuous MESI was initiated ($t = 0$) and used to monitor the awake subject for 10 minutes before increasing isoflurane to 2.0\% to induce anesthesia. After one minute of exposure ($t = 11$ minutes), the nose-cone was fully positioned over the subject to ensure targeted delivery of isoflurane. After two minutes of exposure ($t = 12$ minutes), a feedback heating pad (55-7030, Harvard Apparatus, Inc.) was placed beneath the subject to maintain a 37 $^\circ$C body temperature. The now anesthetized subject was continuously monitored with MESI until $t = 30$ minutes, at which point the anesthetized hemodynamic state was defined by acquiring and averaging an additional 50 MESI frames. The subject was then removed from the imaging system and allowed to recover from anesthesia on a heating pad before being transferred back into its cage. Each subject underwent three imaging sessions with at least one full day of recovery between each session.

@@

All data processing was performed using MATLAB (R2021a, MathWorks, Inc.). While head fixation can greatly minimize locomotion-associated brain movement \cite{Dombeck:2007gr}, it does not completely eliminate motion artifacts. The induction of general anesthesia also causes major changes in posture that can result in lateral displacements of the brain relative to the skull. In order to account for these shifts, the open source image registration toolbox \texttt{Elastix} \cite{Klein:2010gr} was used to rigidly align all speckle contrast frames to the baseline awake imagery prior to fitting Eq.~(\ref{eq:mesi}). The resulting aligned $ICT$ frames were then smoothed temporally with a central moving average filter ($k=5$) and rendered to video at 20 frames-per-second (8$\times$ speed). $ICT$ timecourses spanning the entire imaging session were then extracted from the aligned data from both vascular and parenchymal regions of interest (ROIs).

\subsubsection{Calculating Vascular Flux}
In order to calculate the vascular flux \cite{Kazmi:2015du}, the diameter of vessels ($d_{vessel}$) were estimated using the 5 ms exposure speckle contrast frames. Cross-sectional profiles were extracted across each vessel of interest at every timepoint and fitted to a Gaussian function in order to calculate the full-width at half-maximum (FWHM) of the distribution over time. Because LSCI is most sensitive to moving scatterers, i.e. erythrocytes, this method of estimating $d_{vessel}$ is a measure of the inner diameter of the vessel. The vascular flux was then computed by scaling the $ICT$ by $d_{vessel}$ at each timepoint. The percent change in vessel diameter and vascular flux were calculated using the average of the entire awake state for the baseline.

\subsubsection{Calculating Blood Flow Rise Time}
To examine the dynamics of isoflurane-induced flow changes, the rise time in measured blood flow was calculated. Rise time was defined as the time elapsed following the start of anesthesia for flow values to increase by 50\% relative to the awake baseline. A spatially-resolved blood flow rise time map based on $ICT$ was generated for Subject 1 by performing this calculation pixel-wise across the entire field-of-view. Analogous rise time calculations were performed on the vascular flux and parenchymal tissue perfusion metrics across all subjects and imaging sessions using the previously extracted ROI timecourses.

\subsubsection{Statistical Analysis}
All statistical analyses were also performed using MATLAB. The differences between arterial, venous, and parenchymal measurements were evaluated with one-way ANOVAs using the Bonferroni method for multiple comparisons. $p \leq 0.05$ was considered statistically significant.

@@

The systematic changes caused by general anesthesia with isoflurane in Subject 1 are shown in Fig.~\ref{fig:system}b and Video 1, which highlights the transition from the awake to the anesthetized state. The visible decrease in speckle contrast and corresponding increase in $ICT$ are representative of an increase in CBF caused by the vasodilatory effects of isoflurane. Dilation is seen across much of the surface vasculature, in particular the large central vein and numerous smaller vessels that have become more prominent in the anesthetized imagery as blood flow increased. These large alterations to the vasculature complicated rigid image registration and precluded the generation of relative flow change imagery.

% Figure 1 - System Schematic + Awake vs. Anesthetized Imagery
\begin{figure}
    \includegraphics[width=\textwidth]{Figure1.pdf}
    \caption {
        (a) MESI system schematic with 2$\times$ magnification: A (Standard Iris, ID15, Thorlabs, Inc.), M ({{\O}}1$"$ Broadband Dielectric Mirror, BB1-E02, Thorlabs, Inc.), L1 ($f =$ 50 mm, Steinheil Triplet Achromatic, 67-422, Edmund Optics, Inc.), F (685 $\pm$ 20 nm bandpass filter, S685/40m, Chroma Technology Corp.), and L2 ($f = 100$ mm, Achromatic Doublet, AC254-100-A, Thorlabs, Inc.). (b) Single exposure LSCI (5 ms) and MESI during the awake and anesthetized states in Subject 1 (Scale bar = 1 mm). See Video 1 (MPEG-4, 29.8 MB) for the transition between each state during the induction of general anesthesia.
    }
    \label{fig:system}
\end{figure}

The temporal dynamics of the awake-to-anesthetized transition can be seen in Fig.~\ref{fig:timecourse}, where the change in blood flow within two veins (R1 and R2) and one parenchyma region (R3) were analyzed during the extended two-hour imaging session in Subject 1. The cross sections used to estimate $d_{vessel}$ are depicted in Fig.~\ref{fig:timecourse}a with the resulting changes in vessel diameter shown in Fig.~\ref{fig:timecourse}b. The width of the two vessels remained steady throughout the entirety of the awake measurements but increased under anesthesia, with the larger vessel (R1) growing by over 50\% after an hour of exposure. This vasodilation was also accompanied by an increase in blood flow as measured by the percent change in vascular flux and parenchymal tissue perfusion as shown in Fig.~\ref{fig:timecourse}c. The vascular flux in R1 and R2 increased by 270\% and 144\%, respectively, while the tissue perfusion in R3 grew by 141\%. The large deviation shortly after $t = 80$ minutes was caused by the subject briefly experiencing breathing difficulties that required the anesthesia be reduced to 1.5\% isoflurane for several minutes. The breathing abnormality and change in anesthesia dosage had minimal effect on the vessel size. Repeated measurements in the subject (Figs.~\ref{fig:timecourse}d-f) illustrate the stability of day-to-day MESI estimates of blood flow as well as the consistency of the hemodynamic response to isoflurane anesthesia.

@@

The percent changes in vessel diameter and blood flow between the awake and anesthetized states were aggregated across all subjects and imaging sessions as shown in Fig.~\ref{fig:aggregated}. The anesthetized measurements for Subject 1, which underwent the extending imaging sessions, were taken at $t \approx 80$ minutes to match the timing of the other subjects (20 minutes after the start of anesthesia). A total of 108 measurements were included in the analysis (33 arteries, 29 veins, 46 parenchyma regions). Similar to the results seen in Fig.~\ref{fig:timecourse}, the vasodilatory effects of isoflurane induced large changes in the cortical hemodynamics. On average, vessel width increased by 14.1\%, with arteries increasing by 15.0\% and veins increasing by 13.0\% (Fig.~\ref{fig:aggregated}a). There was no significant difference in the percent change in width between arteries and veins ($p = 0.61$). Vascular flux increased, on average, by 96.0\%, with arteries increasing by 81.7\% and veins increasing by 112.2\% (Fig.~\ref{fig:aggregated}b). Parenchymal tissue perfusion increased by 84.7\%. While the ANOVA was significant ($p = 0.041$), the multiple pairwise comparisons between arteries, veins, and parenchyma regions were all non-significant ($p > 0.05$).

% Figure 3 - Aggregated Awake to Anesthetized Results
\begin{figure}
    \includegraphics[width=\textwidth]{Figure3.pdf}
    \caption {
        (a) Percent change in vessel diameter for arteries and veins between the awake and anesthetized states ($p > 0.05$). (b) Percent change in arterial and venous vascular flux and parenchymal tissue perfusion between the awake and anesthetized states ($p > 0.05$). Measurements in both charts aggregated across all subjects and all imaging sessions.
    }
    \label{fig:aggregated}
\end{figure}

The dynamics of the isoflurane-induced flow changes in Subject 1 can be seen in the blood flow rise time map in Fig.~\ref{fig:risetime}a. $ICT$ values across the entire field-of-view increased by at least 50\% in just over one minute following the start of anesthesia. The vasculature, including both arteries and veins, appears to change flow faster than parenchyma regions. However, upon aggregating ROI measurements across all subjects and imaging sessions (Fig.~\ref{fig:risetime}b), arteries have significantly faster rise times ($40.5$ seconds) compared to veins ($50.7$ seconds, $p = \num{3.7e-6}$) or parenchyma regions ($46.1$ seconds, $p = 0.006$). Parenchymal regions also have significantly faster rise times than veins ($p = 0.043$).

% Figure 4 - Hemodynamic Rise Time
\begin{figure}
    \includegraphics[width=\textwidth]{Figure4.pdf}
    \caption {
        (a) Blood flow rise time map depicting the time elapsed following the start of anesthesia for $ICT$ to increase by 50\% relative to the awake baseline in Subject 1 (Scale bar = 1 mm). (b) Time for arterial and venous vascular flux and parenchymal tissue perfusion to increase by 50\% relative to the awake baseline (***$p = \num{3.7e-6}$, **$p = 0.006$, *$p = 0.043$). Measurements aggregated across all subjects and all imaging sessions.
    }
    \label{fig:risetime}
\end{figure}

@@

The systemic effects of isoflurane undermine the reliability of neurophysiological and behavioral studies because anesthesia is a fundamentally unnatural physiological state \cite{Gao:2017tw}. Imaging blood flow on the cortical surface with MESI during the induction of anesthesia directly visualizes and quantifies these large hemodynamic changes. On average, the vasodilation caused by isoflurane increased vessel diameter by 14.1\% across all trials, with arteries increasing by 15.0\% and veins increasing by 13.0\% (Fig.~\ref{fig:aggregated}a). Direct microscope measurements in fentanyl- and nitrous oxide-anesthetized rats documented a 17\% increase in arteriolar diameter and 6\% increase in venule diameter after the topical application of isoflurane \cite{Koenig:1994rn}. Similar measurements in pentobarbital-anesthetized dogs observed 10-28\% increases in the diameter of small arterioles following the inhalation of isoflurane \cite{Iida:1998th}. Both studies found that isoflurane dilated arterioles in a concentration-dependent manner. More recent measurements in mice using optical coherence tomography (OCT) angiography found that isoflurane increased artery diameter by 12-55\% and vein diameter by 14-22\%, depending on vessel branch order \cite{Rakymzhan.2021}.

Much larger changes were measured in surface vessel blood flow, with vascular flux increasing on average by 96.0\% across all trials (Fig.~\ref{fig:aggregated}b). This suggests that increases in blood flow velocity rather than vasodilation are largely responsible for the measured change in flow. Because few imaging modalities are capable of measuring the dynamics of blood flow with sufficient spatial resolution to distinguish individual vessels, there has been limited work on this topic in awake animals. One study using laser Doppler flowmetry found only non-significant 18\% increases in both CBF and red blood cell (RBC) velocity in the barrel cortex after mice were anesthetized with isoflurane \cite{Takuwa:2012ee}. However, because large blood vessels were avoided, the laser Doppler measurements only sampled smaller subsurface vasculature that may exhibit different responses to isoflurane than the larger surface vasculature imaged in this study. The vascular flux metric calculated from the speckle correlation time is also not analogous to either the CBF or RBC velocity measured by laser Doppler flowmetry, so the results are not directly comparable \cite{Kazmi:2015du}. Another more recent study using Doppler OCT measured a 55\% increase in volumetric blood flow in mice under isoflurane \cite{Rakymzhan.2021}. Unlike the results presented above, this change was largely driven by vasodilation rather than an increase in blood flow velocity.

Within the parenchyma, tissue perfusion increased on average by 84.7\% across all trials after isoflurane exposure (Fig.~\ref{fig:aggregated}b). The narrower distribution of values compared to the vascular measurements is likely a byproduct of the more homogenous structure of the unresolvable capillaries within the parenchyma. fMRI measurements in isoflurane-anesthetized rats have reported $\sim$50\% increases in global CBF with the cerebral cortex experiencing the smallest regional increase (20\%) compared to the awake state \cite{Sicard:2003dj}. However, these numbers are not directly comparable because fMRI samples a much larger volume of the brain than MESI, which has both a smaller field-of-view and only penetrates several hundred microns into the cortex.

While LSCI has been utilized extensively with awake imaging \cite{Takuwa:2011jr,Seto:2014ga,Lu:2017bo,Balbi:2017cj,Sunil:2020ac}, this is one of the first uses of MESI in an awake animal model \cite{He:2020}. Traditional single-exposure LSCI would likely have underestimated the magnitude of the flow changes caused by isoflurane because individual exposure times are only sensitive to a narrow range of flow rates \cite{Parthasarathy:2008el}. Failing to account for the increase in multiple dynamic scattering events caused by vasodilation would also have resulted in an underestimation of the flow changes in the surface vasculature. Using the raw $ICT$ value instead of the diameter-scaled vascular flux metric in Fig.~\ref{fig:timecourse}c would have measured only a 141\% increase in the large vessel instead of 270\%. The repeatability of the day-to-day measurements shown in Figs.~\ref{fig:timecourse}d-f further demonstrate the robustness of MESI as a chronic blood flow imaging technique \cite{Kazmi:2013hp}, even when observing a fully-conscious, moving animal.

@@

The effects of isoflurane fundamentally undermine the imaging of cortical hemodynamics in the normal physiological state. Blood flow imaging techniques such as LSCI, laser Doppler flowmetry, and fMRI require greater sensitivity because of the suppression of the hemodynamic response \cite{Takuwa:2012ee} and measure a delayed neurovascular coupling \cite{Pisauro:2013cx}. Anatomical imaging with two-photon microscopy only captures heavily-dilated vasculature unrepresentative of the awake resting state \cite{Lyons:2016bd}. The consistency and reliability of all imaging techniques can suffer from the effects of prolonged isoflurane exposure, which can cause continual vasodilation as seen in Fig.~\ref{fig:timecourse}b. This could cause extended imaging sessions to document disparate anatomies and physiologies between the beginning and conclusion of an experiment.
